---
title: "Project 1"
output:
  rmarkdown::github_document:
    toc: true
    toc_depth: 2
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes  
           
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#rmarkdown::render("project1.Rmd", params = "ask")
```

```{r echo = FALSE, message = FALSE}
#Connecting to NHL records API

# Install and packages.

library(httr)
library(jsonlite)
library(dplyr)
library(ggplot2)
library(DT)
library(knitr)
library(tidyverse)

#connecting to the data (5 in total)

# /franchise (Returns id, firstSeasonId and lastSeasonId and name of every team in the history of the NHL)
base_url1 <- "https://records.nhl.com/site/api"
tab_name1 <- "franchise"


# Paste the pieces together like so:
full_url1 <- paste0(base_url1, "/",tab_name1,"/")

get_poke1 <- GET(full_url1)
txt_poke1 <- content(get_poke1, "text")
json_poke1 <- fromJSON(txt_poke1, flatten=TRUE)
df_poke1 <- as.data.frame(json_poke1)

#----------------------------------------

# /franchise-team-totals (Returns Total stats for every franchise (ex roadTies, roadWins, etc))
base_url2 <- "https://records.nhl.com/site/api"
tab_name2 <- "franchise-team-totals"


# Paste the pieces together like so:
full_url2 <- paste0(base_url2, "/",tab_name2,"/")

get_poke2 <- GET(full_url2)
txt_poke2 <- content(get_poke2, "text")
json_poke2 <- fromJSON(txt_poke2, flatten=TRUE)
df_poke2 <- as.data.frame(json_poke2)

#----------------------------------------

# /site/api/franchise-season-records?cayenneExp=franchiseId=ID (Drill-down into season records for a specific franchise)

full_url3 <- "https://records.nhl.com/site/api/franchise-season-records?cayenneExp=franchiseId=31"

get_poke3 <- GET(full_url3)
txt_poke3 <- content(get_poke3, "text")
json_poke3 <- fromJSON(txt_poke3, flatten=TRUE)
df_poke3 <- as.data.frame(json_poke3)

#----------------------------------------

# /franchise-goalie-records?cayenneExp=franchiseId=ID (Goalie records for the specified franchise)

full_url4 <- "https://records.nhl.com/site/api/franchise-goalie-records?cayenneExp=franchiseId=1"

get_poke4 <- GET(full_url4)
txt_poke4 <- content(get_poke4, "text")
json_poke4 <- fromJSON(txt_poke4, flatten=TRUE)
df_poke4 <- as.data.frame(json_poke4)

#----------------------------------------

# /franchise-skater-records?cayenneExp=franchiseId=ID (Skater records, same interaction as goalie endpoint)

full_url5 <- "https://records.nhl.com/site/api/franchise-skater-records?cayenneExp=franchiseId=1"

get_poke5 <- GET(full_url5)
txt_poke5 <- content(get_poke5, "text")
json_poke5 <- fromJSON(txt_poke5, flatten=TRUE)
df_poke5 <- as.data.frame(json_poke5)

```


```{r echo = FALSE, message = FALSE}

#Connecting to NHL records API (8 total)

# ?expand=team.roster Shows roster of active players for the specified team
full_url6 <- "https://statsapi.web.nhl.com/api/v1/teams?expand=team.roster"
#tab_name6 <- "expand=team.roster"


# Paste the pieces together like so:
#full_url6 <- paste0(base_url6, "?",tab_name6,"/")

get_poke6 <- GET(full_url6)
txt_poke6 <- content(get_poke6, "text")
json_poke6 <- fromJSON(txt_poke6, flatten=TRUE)
df_poke6 <- as.data.frame(json_poke6)

#----------------------------------------

# ?expand=person.names Same as above, but gives less info.
full_url7 <- "https://statsapi.web.nhl.com/api/v1/teams?expand=person.names"
#tab_name7 <- "expand=person.names"


# Paste the pieces together like so:
#full_url7 <- paste0(base_url7, "?",tab_name7,"/")

get_poke7 <- GET(full_url7)
txt_poke7 <- content(get_poke7, "text")
json_poke7 <- fromJSON(txt_poke7, flatten=TRUE)
df_poke7 <- as.data.frame(json_poke7)

#----------------------------------------

# ?expand=team.schedule.next Returns details of the upcoming game for a team
full_url8 <- "https://statsapi.web.nhl.com/api/v1/teams?expand=team.schedule.next"
#tab_name8 <- "expand=team.schedule.next"


# Paste the pieces together like so:
#full_url8 <- paste0(base_url8, "?",tab_name8,"/")

get_poke8 <- GET(full_url8)
txt_poke8 <- content(get_poke8, "text")
json_poke8 <- fromJSON(txt_poke8, flatten=TRUE)
df_poke8 <- as.data.frame(json_poke8)

#----------------------------------------

# ?expand=team.schedule.previous Same as above but for the last game played
full_url9 <- "https://statsapi.web.nhl.com/api/v1/teams?expand=team.schedule.previous"
#tab_name9 <- "expand=team.schedule.previous"


# Paste the pieces together like so:
#full_url9 <- paste0(base_url9, "?",tab_name9,"/")

get_poke9 <- GET(full_url9)
txt_poke9 <- content(get_poke9, "text")
json_poke9 <- fromJSON(txt_poke9, flatten=TRUE)
df_poke9 <- as.data.frame(json_poke9)

#----------------------------------------

# ?expand=team.stats Returns the teams stats for the season
full_url10 <- "https://statsapi.web.nhl.com/api/v1/teams?expand=team.stats"
#tab_name10 <- "expand=team.stats"


# Paste the pieces together like so:
#full_url10 <- paste0(base_url10, "?",tab_name10,"/")

get_poke10 <- GET(full_url10)
txt_poke10 <- content(get_poke10, "text")
json_poke10 <- fromJSON(txt_poke10, flatten=TRUE)
df_poke10 <- as.data.frame(json_poke10)

#----------------------------------------

# ?expand=team.roster&season=20142015 Adding the season identifier shows the roster for that season
full_url11 <- "https://statsapi.web.nhl.com/api/v1/teams?expand=team.roster&season=20142015"

get_poke11 <- GET(full_url11)
txt_poke11 <- content(get_poke11, "text")
json_poke11 <- fromJSON(txt_poke11, flatten=TRUE)
df_poke11 <- as.data.frame(json_poke11)

#----------------------------------------

# ?teamId=4,5,29 Can string team id together to get multiple teams
full_url12 <- "https://statsapi.web.nhl.com/api/v1/teams?teamId=4,5,29"

get_poke12 <- GET(full_url12)
txt_poke12 <- content(get_poke12, "text")
json_poke12 <- fromJSON(txt_poke12, flatten=TRUE)
df_poke12 <- as.data.frame(json_poke12)

```


```{r echo = FALSE}
# wrapper function to choose any endpoint
wrapperfunct <- function(x) {
  if (franchise) {
  return(df_poke1)

    } else if (franchise-team-totals) {
  return(df_poke2)

      } else if (franchise-season-records) {
  return(df_poke3)

        } else if (franchise-goalie-records) {
  return(df_poke4)

          } else if (franchise-skater-records) {
  return(df_poke5) 

            } else if (team.roster) {
  return(df_poke6)

              } else if (person.names) {
  return(df_poke7)

                } else if (team.schedule.next) {
  return(df_poke8)

                  } else if (team.schedule.previous) {
  return(df_poke9) 
             
                     } else if (team.stats) {
  return(df_poke10)
               
                        } else if (team.roster&season) {
  return(df_poke11)
                 
                           } else if (teamId) {
  return(df_poke12) 
} 
}

```


## NHL

##### What is NHL

NHL stands for _National Hockey League_ and currently has 31 teams. My knowledge on the NHL is very small so first I wanted to know the teams with the most wins and the most losses. To view the updated season, check [this website](https://www.nhl.com).

## Top Total Wins/Losses

From looking at the bar charts, Montreal Canadiens has the most wins. An interesting analysis I found is that Boston Bruins have the 2nd most total wins but they have the 5th most losses. This makes me assume the franchise was around longer than other teams as it's in 2nd place for most games played. If I were to choose to watch a team, it would be Montreal Canadiens as they aren't on the Top 5 Losses list. This makes me believe they are the most successful franchise.

```{r echo=FALSE, message = FALSE}

# Top 5 Hockey teams with most wins
bestteams  <- group_by(df_poke2, data.teamName)
bestteams  <-  arrange(bestteams, desc(data.wins))
bestteams <- bestteams[1:5,]

j <- ggplot(bestteams, aes(x = reorder(data.teamName, -data.wins), y = data.wins))
j + geom_bar(stat = "identity")  +
  labs(x = "Hockey Teams", y = "Total Wins") +
     geom_text(aes(label=data.wins), position=position_dodge(width=0.9), vjust=-0.25) +
        ggtitle("Top 5 Hockey Teams with Most Wins")

# Top 5 Hockey teams with most losses
worstteams  <- group_by(df_poke2, data.teamName)
worstteams  <-  arrange(worstteams, desc(data.losses))
worstteams <- worstteams[1:5,]

k <- ggplot(worstteams, aes(x = reorder(data.teamName, -data.losses), y = data.losses))
k + geom_bar(stat = "identity")  +
  labs(x = "Hockey Teams", y = "Total Losses") +
     geom_text(aes(label=data.losses), position=position_dodge(width=0.9), vjust=-0.25) +
        ggtitle("Top 5 Hockey Teams with Most Losses")

# Top 5 Hockey teams Games Played
gpteams  <- group_by(df_poke2, data.teamName)
gpteams  <-  arrange(gpteams, desc(data.gamesPlayed))
gpteams <- gpteams[1:5,]

l <- ggplot(gpteams, aes(x = reorder(data.teamName, -data.gamesPlayed), y = data.gamesPlayed))
l + geom_bar(stat = "identity")  +
  labs(x = "Hockey Teams", y = "Games Played") +
     geom_text(aes(label=data.gamesPlayed), position=position_dodge(width=0.9), vjust=-0.25) +
        ggtitle("Top 5 Hockey Teams Games Played")

```

## Tampa Bay Lightning

I currently reside in Tampa, Florida; so I did an analysis on the _Tampa Bay Lightning_.Here is the roster for the current season.The roster has a total of 27 active players. 

__Active Roster__
```{r echo=FALSE, message = FALSE}

dfpokeTampa <-  filter(df_poke6, teams.name == "Tampa Bay Lightning")
dfpokeTampa <- as.data.frame(dfpokeTampa$teams.roster.roster)
dfpokeTampa <- select(dfpokeTampa, jerseyNumber, person.fullName, position.name, position.type)
dfpokeTampa3 <- rename(dfpokeTampa, "Jersey Number" = "jerseyNumber")
dfpokeTampa3 <- rename(dfpokeTampa3, "Full Name" = "person.fullName")
dfpokeTampa3 <- rename(dfpokeTampa3, "Position Name" = "position.name")
dfpokeTampa3 <- rename(dfpokeTampa3, "Position Type" = "position.type")
knitr::kable(dfpokeTampa3)

```


The table and histogram below shows the count for each position on the Tampa Bay Lightning. Center is the highest count so I assume this must be the most important position on the team or they are required more during the game. There's only 3 goalies and this makes sense as you only need one goalie while the game is going on.

```{r echo=FALSE, message = FALSE, warning = FALSE}

#Count for each position
counttable <- table(dfpokeTampa$position.name)

#table
kable(counttable, col.names = c("Position", "Count"),caption = "Tampa Bay Lightning Position Name Count")

#histogram
a <- ggplot(dfpokeTampa, aes(position.name))
a + geom_histogram(binwidth = 5, stat = "count")

```




The table below shows the percentages for win, losses, and ot for the Tampa Bay Lightning for the recent season. Since it is above 0.50, this means the team is on the right track as they are winning more games than losing or getting ot. To view their updated season, check [this website](https://www.nhl.com/lightning).

```{r echo=FALSE, message = FALSE, warning = FALSE}

dfpokeTampa2 <-  filter(df_poke10, teams.name == "Tampa Bay Lightning")
dfpokeTampa2 <- as.data.frame(dfpokeTampa2$teams.teamStats)
dfpokeTampa2 <- as.data.frame(dfpokeTampa2$splits)

# create 2 new variables for win percentage and loss percentage


dfpokeTampawinloss <- transmute(dfpokeTampa2,
  Win_Percentage = as.integer(stat.wins) / stat.gamesPlayed,
  Loss_Percentage = as.integer(stat.losses) / stat.gamesPlayed,
  OT_Percentage = as.integer(stat.ot) / stat.gamesPlayed
)
 
kable(dfpokeTampawinloss[1,], caption = "Tampa Bay Lightning Win/Loss Percentage")

```

## Tampa Bay vs Atlanta Division Teams

Tampa Bay Lightning is in the Atlantic Division along with 7 other teams. Let's see the Atlantic Division stats compared to each other.The median seems to be around 33 and out of a season of 70 games, the median for Atlantic Division could use some improvement overall.There's a max of 44 by Boston Bruins which means they're doing well. Detroit has 17 wins which makes them a very weak team.

```{r echo=FALSE, message = FALSE, warning = FALSE}

dfpokeTampaBay2 <- dfpokeTampa2[1,]

#Boston Bruins Stats
dfpokeBoston2 <-  filter(df_poke10, teams.name == "Boston Bruins")
dfpokeBoston2 <- as.data.frame(dfpokeBoston2$teams.teamStats)
dfpokeBoston2 <- as.data.frame(dfpokeBoston2$splits)
dfpokeBoston2 <- dfpokeBoston2[1,]

#Toronto Maple Leafs
dfpokeToronto2 <-  filter(df_poke10, teams.name == "Toronto Maple Leafs")
dfpokeToronto2 <- as.data.frame(dfpokeToronto2$teams.teamStats)
dfpokeToronto2 <- as.data.frame(dfpokeToronto2$splits)
dfpokeToronto2 <- dfpokeToronto2[1,]

#Florida Panthers
dfpokeFL2 <-  filter(df_poke10, teams.name == "Florida Panthers")
dfpokeFL2 <- as.data.frame(dfpokeFL2$teams.teamStats)
dfpokeFL2 <- as.data.frame(dfpokeFL2$splits)
dfpokeFL2 <- dfpokeFL2[1,]

#Montreal Canadiens
dfpokeMontreal2 <-  filter(df_poke10, teams.name == "Montréal Canadiens")
dfpokeMontreal2 <- as.data.frame(dfpokeMontreal2$teams.teamStats)
dfpokeMontreal2 <- as.data.frame(dfpokeMontreal2$splits)
dfpokeMontreal2 <- dfpokeMontreal2[1,]

#Buffalo Sabres
dfpokeBuffalo2 <-  filter(df_poke10, teams.name == "Buffalo Sabres")
dfpokeBuffalo2 <- as.data.frame(dfpokeBuffalo2$teams.teamStats)
dfpokeBuffalo2 <- as.data.frame(dfpokeBuffalo2$splits)
dfpokeBuffalo2 <- dfpokeBuffalo2[1,]

#Ottawa Senators
dfpokeOttawa2 <-  filter(df_poke10, teams.name == "Ottawa Senators")
dfpokeOttawa2 <- as.data.frame(dfpokeOttawa2$teams.teamStats)
dfpokeOttawa2 <- as.data.frame(dfpokeOttawa2$splits)
dfpokeOttawa2 <- dfpokeOttawa2[1,]

#Detroit Red Wings
dfpokeDetroit2 <-  filter(df_poke10, teams.name == "Detroit Red Wings")
dfpokeDetroit2 <- as.data.frame(dfpokeDetroit2$teams.teamStats)
dfpokeDetroit2 <- as.data.frame(dfpokeDetroit2$splits)
dfpokeDetroit2 <- dfpokeDetroit2[1,]

#Union all Atlantic Division teams together
atldivision <- union(dfpokeBoston2, dfpokeTampaBay2)

atldivision <- union(atldivision, dfpokeToronto2)
atldivision <- union(atldivision, dfpokeFL2)
atldivision <- union(atldivision, dfpokeMontreal2)
atldivision <- union(atldivision, dfpokeBuffalo2)
atldivision <- union(atldivision, dfpokeOttawa2)
atldivision <- union(atldivision, dfpokeDetroit2)

atldivision <- mutate(atldivision,
                      season = "2019-2020")

boxplot(as.integer(stat.wins)~season,
        data = atldivision,
        main="Atlantic Division Wins",
        xlab="2019-2020 Season",
        ylab="Wins",
        col="orange",
        border="brown"
)


```

From the plot below, we can see a trend that the higher the percentage is that you out shoot the opponent, the higher wins you will get. 

```{r  echo=FALSE, message = FALSE, warning = FALSE}
#Join to get team abbreviation

#Create Table for Team Id and Abbreviation
df_poke2abrv <- select(df_poke2, data.teamId, data.triCode)
# Delete Duplicate Rows
todelete <- seq(1, nrow(df_poke2abrv), 2)
df_poke2abrv <- df_poke2abrv[ todelete ,]
df_poke2abrv <- rename(df_poke2abrv, "team.id" = "data.teamId")

#Join
atldivision <- left_join(atldivision, df_poke2abrv, by = "team.id")
atldivision <- rename(atldivision, "Tri Code" = "data.triCode")


atldivision$team.name <- as.factor(atldivision$team.name)
f <- ggplot(atldivision, aes(stat.winOutshootOpp, stat.wins, colour=`Tri Code`))
f + geom_point() +
  xlab("Win % By Out Shooting Opponent") +
  ylab("Wins") +
  ggtitle("Wins by Out Shooting Opponent") + 
  scale_fill_discrete(name = "Teams")  +     
   geom_smooth(method = 'lm', color = 'green')
```

